---
title: "clustering_invest_Strategy_validation"
output: html_document
date: "2024-10-01"
editor_options: 
  markdown: 
    wrap: 80
params:
  is_load: FALSE  
  is_prod: FALSE
  clust_trial: "gmm_prod_exper1_1"
  clust_trial2: "gmm_prod_exper1_1-simp_prod_exper1_3"
  securities: "c('SPX_ETF')"
  no_cores: 16
---

# Task Overview

-   This document details the steps for backtesting a portfolio based on
    clustering model results. The portfolio rebalances at a regular frequency
    (e.g., weekly or monthly).

-   An investor utilizes the market situation determined by the clustering model
    on the model calculation date and performs rebalancing on the trade date,
    which is the day after the model calculation.

-   The investor has insight into the performance of multiple financial
    instruments from past data from periods with the same market conditions as
    the model execution date, classified by the clustering model. Based on this
    insight, the investor sells all current holdings and buys a selection of
    financial instruments that performed well during those past periods,
    rebalancing the portfolio according to the model’s recommendations. This
    process is repeated following the regular rebalancing schedule.

-   Referring to the past data from periods with the same market conditions as
    the model calculation date is recognized as in-sample results, while the
    investment results on the trade dates are recognized as out-of-sample
    results.

# Preset

## Clean Up the R Environment

-   Before running the clustering investment strategy, it is important to clean
    the R environment to avoid any conflicts with previously loaded data or
    objects. This ensures that only the necessary variables are available,
    preventing unexpected behavior during execution.

```{r}
if(exists("params")){
  keep_vars <- c("params", names(params))
  rm(list = setdiff(ls(), keep_vars))
}
```

## Loading Parameters from YAML Header

-   The following parameters are loaded from the YAML header. These parameters
    control whether to load existing clustering results or perform a new
    analysis, and specify the trial configurations for the investment strategy.

    -   `is_load`: Indicates whether to load previously generated results
        (`TRUE`) or run a new analysis (`FALSE`).
    -   `is_prod`: Indicates if the script is running in production mode. If
        `TRUE`, results are saved in a designated location.
    -   `clust_trial`: Defines a trial for the clustering analysis.
    -   `clust_trial2`: Defines a trial for the combination of a clustering
        analysis and a universe (a set of financial instruments).
    -   `no_cores`: Specifies the number of CPU cores to use for parallel
        processing. This helps optimize the performance of the clustering
        analysis, especially when dealing with large datasets.
    -  `securities`:Used in the create_instruments_return_data function to generate 
    daily returns and total return indices for both long and short positions on these securities.

```{r}
if(exists("params")){
  is_load = params$is_load
  is_prod = params$is_prod
  clust_trial = params$clust_trial 
  clust_trial2 = params$clust_trial2 
  no_cores = params$no_cores
  securities = eval(parse(text = params$securities))
}
```

## Required Libraries

-   The following libraries are required for this analysis.

```{r}
library(tidyverse)
library(conflicted)
library(here)
library(R6)
library(reticulate)
library(zoo)
library(RcppPerfCalc)
library(parallel)
library(fst)

conflicts_prefer(stats::dist)
conflicts_prefer(dplyr::lag)
conflicts_prefer(purrr::map)
conflicts_prefer(dplyr::filter)

```

## Defined Functions and File Path

-   This section loads custom-defined functions from various subfolders, and
    sets the folder paths for loading R scripts and additional utility
    functions.

```{r warning=FALSE}
sim_data_path  = paste0(here(""), "/sim_data")
rcode_path     = paste0(here(""), "/R")
data_save_path = paste0(sim_data_path, "/data/") 
data_save_file_path = paste0(data_save_path, "eq_tri_tbl.rds")
# Load custom R functions from various directories
for (subfolder in c("cluster", "backtest", "utils", "finance")) {
  list.files(here("R", subfolder), "\\.R$", full.names = TRUE) %>% 
    lapply(source)
  list.dirs(here("R", subfolder), recursive = FALSE) %>% 
    list.files("\\.R$", full.names = TRUE) %>% 
    lapply(source)
}
source(paste0(rcode_path, "/run_setup.R"))
```

# Preparation

## Calculate price return index

-   Since we cannot retrieve dividend information for stock indices (e.g., NKY,
    SPX) via Yahoo Finance, calculating a Total Return Index (TRI) is not
    feasible.

-   On the other hand, stocks and ETFs include dividend information in Yahoo
    Finance. In this report, I include data for Toyota and the SPX ETF, but not
    for NKY and TPX ETFs, which lack sufficient historical records.

-   Recommendation: If you require a Total Return Index (TRI) for NKY and TPX,
    you will need to manually source or collect dividend information from
    reliable data providers and adjust the index calculations accordingly.


```{r}

mkt_start_dt = as.Date("2000-01-01")
mkt_data_tbl = create_market_data_for_clustering2(start_dt = mkt_start_dt)
return_result_tbl = create_instruments_return_data(mkt_data_tbl, securities)
eq_ror_tbl = return_result_tbl$eq_ror_tbl
eq_tri_tbl = return_result_tbl$eq_tri_tbl
dir.create(data_save_path, FALSE)
eq_ror_tbl %>% saveRDS(paste0(data_save_path, "eq_ror_tbl.rds"))
eq_tri_tbl %>% saveRDS(paste0(data_save_path, "eq_tri_tbl.rds"))

```

## Set up clust_handler and clust_handler2

-   We use the `ParameterResultHandler` class to manage parameter sets and their
    corresponding results. `clust_handler` is used similarly to how it is in
    `clustering_model_generation.Rmd`, while `clust_handler2` manages the
    combination of clustering analysis parameters and universe information,
    along with their corresponding backtest results.

```{r}
clust_save_path = file.path(sim_data_path, "clust") %+% "/"
clust_handler = ParameterResultHandler$new(clust_save_path)
clust_handler2 = clust_handler$create_subdir_obj("expand", primary_keys = "seq_id2")
# handler = ParameterResultHandler$new(file.path(sim_data_path, "tsd") %+% "/", is_old = F)
```

## Loading Clustering Results

-   The following code loads the parameters and results for the specified
    clustering trial.

-   Notuce: You should generate clustering results about `clust_trial` in
    clustering_model_generation.Rmd. The objects related to the clustering
    models are as follows:

-   `clust_param_tbl`: Loads the parameter tibble for the trial defined by
    `clust_trial`.

-   `clust_result_tbl`: Loads the clustering results for the trial, with the
    results saved under the name `clust_result_tbl`.

-   `clust_start_dt`: Extracts the start date of the clustering process from the
    parameter table.

```{r}
clust_param_tbl  = clust_handler$load_params(clust_trial)
clust_result_tbl = clust_handler$load_results(result_name = "clust_result_tbl", 
                                              trial = clust_trial)
clust_start_dt   = clust_param_tbl$start_dt[[1]]

```

## Clustering Invest Parameters

-   `clust_all_param_tbl` includes a combination of information from
    `clust_param_tbl` and additional investing parameters generated in
    `run_setup.R`. The `clust_trial2` represents a combination of a universe (a
    set of financial instruments) and the criteria parameters used to select the
    best instruments from past data from periods with the same market conditions
    as the model execution date, classified by by the `clust_trial` model.

```{r}
save_path = paste0(data_save_path, "/eq_tri_tbl.rds")
tri_file_loader = TriFileLoader$new(file_path      = data_save_file_path,
                                    from_file_type = "yf_path")
arg_list = getListFromObjs(clust_handler, tri_file_loader)
tmp = execute_clust_invest_setting_factory_element(
  CLUST_INVEST_SETTING_FACTORY, clust_trial2, arg_list)
clust_all_param_tbl = tmp$clust_all_param_tbl
config_tbl          = tmp$config_tbl
all_tri_tbl         = tmp$tri_tbl

clust_handler2$set_params(clust_trial2, clust_all_param_tbl)
clust_handler2$save_config(clust_trial2, config_tbl)

clust_all_param_tbl = clust_handler2$load_params(clust_trial2)
```

-   The columns other than `clust_param_tbl` are shown below.

    -   `trial`: The name of the clustering analysis trial, referring to
        `clust_trial`. This identifies the specific clustering model
        configuration used in the analysis. In this example:
        `"gmm_prod_exper1_1"`(indicating a Gaussian Mixture Model trial)

    -   `universe`: The set of financial instruments for investment. It defines
        the scope of instruments included in the analysis. In this example:
        `"SPX_ETF_LONG_ONLY"` (meaning a long-only strategy for SPX ETFs)

    -   `criteria`:The selection criteria used to rank the instruments. This
        could be based on metrics like Sharpe ratio or other performance
        indicators. In this example: "shr" (representing Sharpe ratio)

    -   `n_top`:The number of top-performing instruments selected based on the
        specified criteria on past data from periods with the same market
        conditions as the model execution date, classified by the clustering
        model. In this example: `1` (indicating the top 1 instrument is
        selected)

    -   `prob_inf`: This represents a threshold used to compare the membership
        probabilities generated by clustering models like Gaussian Mixture Model
        (GMM). GMM provides a probability indicating how strongly each data
        point belongs to a specific cluster. The `prob_inf` value is the
        threshold used to evaluate whether a data point's membership probability
        in past periods is high enough to be confidently assigned to a cluster.
        Data points that meet this threshold are then used in the insample
        backtest.

        -   For models that do not provide probabilities (e.g., K-means), the
            membership probability is always considered zero. In such cases, all
            data from past periods are used unconditionally in insample the
            backtest.

    -   `ar_inf`: This represents a threshold used to compare the annualized
        return for each instrument in past periods with the same market
        conditions as the model execution date, classified by the clustering
        model. Instruments will be removed if their annualized return is lower
        than this value. In this example: `NA` (no AR filtering applied). #CHECK

    -   `reb_freq`:The rebalancing frequency, which determines how often the
        portfolio is rebalanced.In this example: `"1w"` (indicating weekly
        rebalancing), `"2w"` (bi-weekly), `"eom"` (end of month)

    -   `wd_no`:Likely represents the number of weekdays associated with the
        rebalance schedule. If this value is 3 and reb_freq = "1w", the
        rebalance schedules in 2024/01 are 1/10, 1/17. 1/24 and 1/31(since 1/3
        is a Japan Holiday). In this context, Monday is represented as day 1

    -   `codes`: A list of instruments traded in the backtest. In the backtest,
        these instruments are filtered by `ar_inf` and `prob_inf`, and selected
        based on the `criteria`. The top `n_top` instruments are chosen
        according to the defined criteria.In this Example: c("SPX_ETF_TRI_LONG",
        "ZERO" )

    -   `trial2`:Identifies `clust_trial2`.In this example:
        `"gmm_prod_exper1_1-simp_exper4_1"`

```{r}
clust_all_param_tbl 

```

## Rebalance Schedule Params and Historical Daily returns

-   A rebalance schedule is a set of dates that includes the calculation of a
    clustering model, opening a position, and closing a position.

-   `w_bds_tbl` is a tibble that contains the holding dates based on Japanese
    holidays.

-   `all_w_bds_plp_tbl` includes daily returns on the dates from `w_bds_tbl` for
    the following financial instruments: JT7203_TRI_LONG, JT7203_TRI_SHORT,
    SPX_ETF_TRI_LONG, SPX_ETF_TRI_SHORT, and ZERO. Their meanings are explained
    below.

    -   "XXX_TRI_LONG" ("XXX_TRI_SHORT") refers to the daily return when holding
        a long (short) position on XXX instrument.

    -   "ZERO" refers to the daily return when no financial instruments are
        held. In other words, it represents the daily return as if the money
        were deposited in a bank account with a zero interest rate.

-   `reb_sche_nest_tbl` represents the rebalance schedules, showing how the
    rebalancing frequency is structured and contains the `reb_freq` and `wd_no`
    values related to clust_all_param_tbl. The `reb_sche` column contains a
    tibble with the rebalance schedule, which corresponds to the `reb_freq` and
    `wd_no` values in the same row. The columns in `reb_sche` are shown below.

    -   `ref_date`: This represents the model calculation date for the
        clustering model.

    -   `h_start`: The start date of the holding period. On this date, positions
        are opened based on the signals generated by the model on the
        `ref_date`. Trades are executed on the next trading day after the
        positions are identified.

    -   `h_end`: The end date of the holding period. Positions are held until
        this date, and all trades are closed by the next trading day to realize
        profits. If positions are held across multiple periods, separate
        `h_start` and `h_end` dates are defined for each period.

```{r}
start_dt  = config_tbl$start_dt[[1]]
end_dt    = config_tbl$end_dt[[1]]
# reb_freqs = config_tbl$reb_freqs[[1]]
# plp_type  = config_tbl$plp_type[[1]]
   
w_bds_tbl =readRDS(data_save_path %+% "w_bds_tbl.rds")
# trade_w_bds_tbl =readRDS(data_save_path %+% "trade_w_bds_tbl.rds")
trade_w_bds_tbl = w_bds_tbl

all_w_bds_plp_tbl = create_d_ror_tbl_from_all_tri_tbl(w_bds_tbl, all_tri_tbl, T) %>% 
  drop_na() 
reb_sche_nest_tbl = create_reb_sche_nest_tbl(w_bds_tbl, config_tbl)
```

-   Additionally, the holding period is extended when the next `ref_date` falls
    on a holiday, based on `w_bds_tbl`. As shown in the example below, for the
    holding period with a `ref_date` of "2000-12-26", the `h_end` is not
    "2001-01-03" but "2001-01-10". This is because "2001-01-03" is a holiday
    according to `w_bds_tbl`.

```{r}
reb_sche_nest_tbl$reb_sche[[1]] %>% 
  filter(ref_date >= "2000-12-26") %>% 
  head(3)
```

# Insample

## Generate setup

-   `clust_result_reb_sche_tbl` integrates the clustering results from
    `clust_result_tbl` with the rebalance schedules in `reb_sche_nest_tbl` and
    is used to align the clustering model's outcomes with the rebalance schedule
    dates. It includes not only the clustering model's parameters and rebalance
    schedules, but also processed in-sample inputs such as `is_clust_tbl`,
    `ref_category`, and `ref_prob`. The prefix 'ref' refers to the model
    calculation date. These columns are explained below:

    -   `is_clust_tbl`: Stores information about how the rebalance schedules and
        the probabilities generated by the clustering model before the model
        calculation date are organized.

    -   `ref_category`: The category assigned to a cluster for each model
        calculation date.

    -   `ref_prob`: The probability of belonging to a cluster for each model
        calculation date.

```{r}
 clust_result_reb_sche_tbl = clust_result_tbl %>%  
    group_nest(clust_num, ref_name, model_name)  %>% 
    expand_grid(reb_sche_nest_tbl) %>% 
    rowwise() %>% 
    mutate(
      data = reb_sche %>% 
        left_join(data %>% select(ref_date, is_clust_tbl, ref_category, ref_prob), 
                  by = join_by(ref_date)) %>%
        drop_na(is_clust_tbl) %>% 
        list()
    ) %>% 
    unnest(cols = c(data)) %>% 
    filter(h_start %in% trade_w_bds_tbl$date) %>% rowwise() %>% 
    mutate(
      is_clust_tbl = is_clust_tbl %>% 
        bind_rows() %>% 
        rename(sim_ref_date = date) %>%
        left_join(reb_sche %>% 
                    rename_with(~ paste0("sim_", .x)), by = join_by(sim_ref_date)) %>%
        drop_na(sim_h_start) %>% 
        filter(sim_h_end < h_start)　%>% 
        list()
    ) %>% 
    select(-reb_sche ) %>% 
    ungroup()
```

-   This chunk executes parallel computations to generate position histories
    with probabilities for each row in `clust_result_reb_sche_tbl`. The
    resulting `clust_is_setup_tbl` is an extension of
    `clust_result_reb_sche_tbl` and contains not only these position histories
    as `sim_pos_dates`, but also in-sample parameters such as `n_top`,
    `prob_inf`, `ar_inf`, and `reb_freq` from `clust_all_param_tbl`.

```{r}
if(!is_load){
  
  system.time({
    temp = create_global_objects_for_plp_from_clust_pos_info_tbl(w_bds_tbl)
    global_f = temp$f
    global_variables = temp$variables
    # Defining Objects in the Global Environment
    environment(global_f) <- .GlobalEnv
    list2env(global_variables, envir = .GlobalEnv)
    environment(execute_parallel_computation_on_splits) <- .GlobalEnv
    
    clust_pos_info_tbl = execute_parallel_computation_on_splits(
      clust_result_reb_sche_tbl, split_function, no_cores = no_cores,
      add_libraries = c("mclust", "here", "R6")) %>% 
      select(-is_clust_tbl) %>%
      group_nest(clust_num, ref_name, model_name, reb_freq, wd_no)
    rm(global_f, global_variables)
   #             ユーザ            システム                経過  
   # 142.050000000000   2.840000000000 738.040000000001 
    gc()
    gc()
  })
  clust_is_setup_tbl = clust_all_param_tbl %>% 
    left_join(clust_pos_info_tbl) %>% 
    unnest(data)
  
  clust_handler2$save_results(clust_trial2, "clust_is_setup_tbl", clust_is_setup_tbl)
}else{
  clust_is_setup_tbl = clust_handler2$load_results("clust_is_setup_tbl", clust_trial2)
  
}

```

## Generate Results

-   This chunk executes parallel computations to:

    -   Make weights by evaluating whether `prob` exceeds `prob_inf` in
        `sim_pos_dates` within `clust_is_setup_tbl`.

    -   Calculate daily returns from weights for past periods that share the
        same market conditions as the model execution date, classified by the
        clustering model.

    -   Calculate in-sample results, specifically determining which financial
        instruments to invest in on the trade date, the day following the model
        calculation. These in-sample results are the final output of the
        parallel computations, stored in `clust_is_result_tbl`, and are
        generated using inputs such as daily returns from those past periods,
        applying rules based on in-sample parameters such as `criteria`,
        `n_top`, `prob_inf`, `ar_inf`, and `reb_freq`.

```{r}

if(!is_load){
  keys_origin_dict = list(code = "ZERO")
  # rules based on in-sample parameters
  rule_param_tbl = clust_all_param_tbl %>% 
    distinct(criteria, n_top, ar_inf) %>% 
    mutate(exec_pct_inf = NA)
  
  tmp = create_global_objects_for_parallel_clust_is_result(
    keys_origin_dict, rule_param_tbl, all_w_bds_plp_tbl,
      
    is_debug = F 
  )
  global_f = tmp$f
  global_variables = tmp$variables


  #グローバル環境でのオブジェクト定義
  environment(global_f) <- .GlobalEnv
  list2env(global_variables, envir = .GlobalEnv)
  environment(execute_parallel_computation_on_splits) <- .GlobalEnv
  # split_function2 = function(x) split_function(x, N = N)
  clust_is_result_tbl = clust_is_setup_tbl %>% 
    mutate(id=row_number()) %>% 
    arrange(desc(ref_date)) %>% 
    execute_parallel_computation_on_splits(
      split_function, no_cores = no_cores, 
      add_libraries = c("here", "R6"))
   rm(global_f, global_variables)
  
  clust_handler2$save_results(clust_trial2, "clust_is_result_tbl", clust_is_result_tbl)
}else{
  clust_is_result_tbl = clust_handler2$load_results(result_name =  "clust_is_result_tbl", trial = clust_trial2)

}

```

# Out-of-Sample

## Generate Results

-   `c_rate` represents the trading cost incurred when executing trades on
    financial instruments based on recommendations from the in-sample results.
    In this document, `c_rate` is set to 5 basis points (bp).

-   Recommendation: The 5bp trading cost is an approximate value for highly
    liquid equities. If you have access to more accurate data from sources like
    Bloomberg or Reuters, please adjust the cost accordingly.

-   The function `create_clust_os_setup_tbl` generates calculates out-of-sample
    results (`clust_os_result_tbl`) based on in-sample results and trading
    schedules. It first expands the backtest schedule from `clust_is_result_tbl`
    and transforms the date range into individual trading days, evenly
    distributing positions across the schedule. Then, it merges the trading
    schedule with the total return index to create a daily return table for all
    instruments. Finally, position weights are calculated and used to compute
    the portfolio's daily performance, which is returned as
    `clust_os_result_tbl`.

```{r}
c_rate = 0.0005

if(!is_load){

  clust_trade_w_bds_tbl = trade_w_bds_tbl %>% 
    filter(date >= clust_start_dt)
  
  # c_rate = if(plp_type=="tsd") 0.0 else c_rate
  clust_os_result_tbl = create_clust_os_setup_tbl(
    clust_is_result_tbl, clust_all_param_tbl, clust_trade_w_bds_tbl, all_tri_tbl, 
    c_rate=c_rate, id_name = "seq_id2")

  clust_handler2$save_results(clust_trial2, "clust_os_result_tbl", clust_os_result_tbl)
}else{
  clust_os_result_tbl = clust_handler2$load_results(result_name =  "clust_os_result_tbl", trial = clust_trial2)
}


```

## Extract Data

-   In this chunk, time series data such as daily returns, total return indices,
    and portfolio performance for the out-of-sample period are extracted from
    `clust_os_result_tbl`.

```{r}
if(!is_load){
  
  clust_os_plp_tbl = clust_os_result_tbl %>% 
    # select(clust_num, ref_name, model_name, codes, d_plp_tbl) %>% 
    select(-data , -d_weight_tbl_s) %>% 
    ungroup() %>% 
    unnest(d_plp_tbl) %>% 
    rename(update_codes = codes)
  clust_handler2$save_results(clust_trial2, "clust_os_plp_tbl", clust_os_plp_tbl)
  
  clust_os_perf_tbl = clust_os_plp_tbl %>% 
    select(seq_id2, update_codes,  date, port)  %>% 
    create_perf_tbl(c("seq_id2", "update_codes"), "port") %>% 
    left_join(clust_all_param_tbl, by = join_by(seq_id2)) %>% 
    select(clust_all_param_tbl %>% colnames(), everything())
  clust_handler2$save_results(clust_trial2, "clust_os_perf_tbl", clust_os_perf_tbl)
  
  clust_os_tri_tbl = clust_os_plp_tbl %>% 
    select(-update_codes) %>% 
    with_cumrets2("date",  "seq_id2", "port")
  clust_handler2$save_results(clust_trial2, "clust_os_tri_tbl", clust_os_tri_tbl)

}



```

```{r}
clust_os_perf_tbl %>% 
  mutate(name = paste0(universe, "_", reb_freq)) %>% 
  bind_rows(eq_ror_tbl %>% 
              filter(date >= clust_start_dt & date <= end_dt)%>% 
              create_perf_tbl2(indeces = "date")) %>% 
  select(name, ar, sd, dd, shr, sor, str)

```
