---
title: "clustering_model_generation"
output: html_document
date: "2024-10-01"
editor_options: 
  markdown: 
    wrap: 80
params:
  is_load: FALSE  
  is_prod: TRUE
  clust_trial: "gmm_prod_exper1_1" 
  no_cores: 16
---

# Task Overview

-   This document details the steps for validating a clustering model used to
    classify non-linear market regimes. The main tasks involve configuring
    market signals and the clustering model itself.

## Market Signal Selection

-   Define the combination of market signals that the model will focus on, such
    as volatility or interest rates.

## Model Selection

-   Choose the appropriate clustering model (e.g., K-means, Gaussian Mixture
    Model).

# Preset

## Clean Up the R Environment

-   Before running the clustering model, it is important to clean the R
    environment to avoid any conflicts with previously loaded data or objects.
    This ensures that only the necessary variables are available, preventing
    unexpected behavior during execution.

-   In the following code, we preserve the `params` object, which contains the
    parameters from the YAML header, and remove all other objects from the
    environment.

```{r}
if(exists("params")){
  keep_vars <- c("params", names(params))
  rm(list = setdiff(ls(), keep_vars))
}
```

## Loading Parameters from YAML Header

-   The following parameters are loaded from the YAML header. These parameters
    control whether to load existing clustering results or perform a new
    analysis, and specify the trial configurations for the investment strategy.

    -   `is_load`: Determines whether to load previously generated clustering
        results (`is_load = TRUE`) or to perform a new clustering analysis
        (`is_load = FALSE`).

    -   `is_prod`: Indicates whether the script is running in production mode.
        If `TRUE`, the generated results are saved in the designated storage
        location. If `FALSE`, the generated results are not saved.

    -   `clust_trial`: The name of the current trial for the clustering
        analysis. This parameter integrates the market signals and the
        clustering model configuration for the trial. The possible values for
        `clust_trial` are defined in `run_setup.R`.

    -   `no_cores`: Specifies the number of CPU cores to use for parallel
        processing. This helps optimize the performance of the clustering
        analysis, especially when dealing with large datasets.

```{r}
if(exists("params")){
  is_load = params$is_load
  is_prod = params$is_prod
  clust_trial = params$clust_trial 
  no_cores = params$no_cores
}
```

## Required Libraries

-   The following libraries are needed for this analysis:

    -   Some functions are conflicted between libraries. Therefore, we use
        `conflicts_prefer` to explicitly specify which functions to use when
        conflicts occur.

```{r}

library(tidyverse)
library(mclust)
library(conflicted)
library(here)
library(R6)
library(zoo)
library(rlang)
library(reticulate)
library(parallel)
library(fst)

conflicts_prefer(stats::dist)
conflicts_prefer(dplyr::lag)
conflicts_prefer(purrr::map)
conflicts_prefer(dplyr::filter)
 
```

## Defined Functions and File Path

-   The following code loads custom-defined functions and sets the folder paths
    for saving generated data and for the R code directory.

    -   `sim_data_path`: Specifies the path where generated data will be saved.

    -   `rcode_path`: Specifies the directory where the R code files are
        located.

```{r}

sim_data_path = paste0(here(""), "/sim_data")
rcode_path = paste0(here(""), "/R")

for (subfolder in c("cluster", "sqlite", "utils", "finance")) {
  list.files(here("R", subfolder), "\\.R$", full.names = TRUE) %>% 
    lapply(source)
  list.dirs(here("R",subfolder), recursive = FALSE) %>% 
    list.files("\\.R$", full.names = TRUE) %>% 
    lapply(source)
}
source(paste0(rcode_path, "/run_setup.R"))
```

# Flow

-   The clustering model is executed for each model execution date within the
    period from `clust_start_dt` to `clust_end_dt`.

-   On each model execution date, the clustering model incrementally
    incorporates new market data from `mkt_start_dt` to `clust_start_dt`. This
    allows the model to adjust to evolving market conditions and refine the
    classification of market regimes as additional data becomes available.

-   This report saves the clustering model parameters in `clust_param_tbl` and
    their corresponding execution results in `clust_result_tbl`.

-   The ParameterResultHandler class manages parameter sets and their
    corresponding results. In this report, the parameter sets are stored in
    param_tbl, and the corresponding results are stored in clust_result_tbl.
    This class saves param_tbl and clust_result_tbl, which are generated during
    the clustering process.

# Preparation

## Market Data

-   This function loads market data from Yahoo Finance using the Python library
    yfinance. The retrieved data will be stored in mkt_data_tbl, a wide-format
    tibble. Please set up an Anaconda environment to run the function or,
    alternatively, prepare your own market data and assign it to mkt_data_tbl.

-   From Yahoo Finance, the function retrieves data for the following symbols:
    USDJPY (Japanese Yen to US Dollar exchange rate), SPX (S&P 500 index), VIX
    (Volatility Index), USGG2YR (US 2-year Treasury yield), and USGG10YR (US
    10-year Treasury yield).

-   The mkt_data_tbl contains the following columns: date, USDJPY, SPX, VIX,
    USGG2YR, and USGG10YR. If you do not use an Anaconda environment, ensure
    that your custom market data contains these columns, with historical values
    that match the date range specified in the function.

```{r warning=FALSE, message = FALSE}
mkt_start_dt = as.Date("2000-01-01")
mkt_data_tbl = create_market_data_for_clustering2(
                                                  start_dt = mkt_start_dt) %>%
  mutate(is_nenmatsu = year(date) != lead(year(date)), 
         is_nenmatsu = replace_na(is_nenmatsu, T)
  )  


```

## Clustering Parameters

-   clust_param_tbl includes the market signals and clustering model parameters
    generated in run_setup.R. The clust_trial represents a combination of the
    market signals and clustering model parameters, allowing you to select
    options such as "GMM_PROD_EXPER1" or "KMEANS_PROD_EXPER1".

```{r}
clust_start_dt = "2002-01-01"
clust_end_dt   = "2023-12-31"

save_path = file.path(sim_data_path, "clust") %+% "/"

clust_config_setting = env_get(CLUST_CONFIG_SETTING, clust_trial)
clust_trial 　　　　 = clust_config_setting$clust_trial
clust_param_tbl      = clust_config_setting$clust_param_tbl

clust_handler = ParameterResultHandler$new(save_path)
clust_handler$set_params(clust_trial, clust_param_tbl)
```

-   The clust_param_tbl is shown in the following chunk, and its columns are
    explained below:

    -   `model_name`: The name of the clustering model being used. In this
        example, the model is "GMM" (Gaussian Mixture Model).

    -   `lib_name`: The name of the library used to implement the clustering
        model. In this example, the library is "Mclust," which is commonly used
        for Gaussian Mixture Model clustering in R.

    -   `start_dt`: The start date for the market data used in the clustering
        process. In this example, the start date is "2002-01-01."

    -   `seed`: The random seed used to ensure reproducibility of the clustering
        results. The seed is set to `123` in this example.

    -   `clust_num`: The number of clusters to be formed in the clustering
        model. In this case, the model is set to form `3` clusters.

    -   `ref_name`: A list of reference market signals or features used as
        inputs for the clustering model. In this example, the reference signals
        are "VIX" (Volatility Index) and "USGG10YR" (US 10-year Treasury yield).

    -   `clust_trial`: The name or identifier for the current trial of the
        clustering process. It helps to distinguish between different experiment
        runs. In this case, the trial is named "gmm_prod_exper1_1."

    -   `comb_ref_name`: A string representing the combination of reference
        market signals used in the clustering model. It is generated from the
        `ref_name` list. In this example, the combination is "VIX\@USGG10YR",
        indicating that both the VIX and USGG10YR signals are used together.

```{r}
clust_param_tbl
```

-   The `clust_param_date_tbl` loads the `clust_param_tbl` for each model
    execution date.

```{r}
date_tbl = mkt_data_tbl %>% 
  filter(date >= clust_start_dt) %>% 
  select(date, is_nenmatsu)

clust_param_date_tbl = clust_param_tbl %>% rowwise() %>% 
  mutate(date_tbl = date_tbl %>% list()) %>% 
  unnest(c(date_tbl )) %>% 
  rename(save_graphs = is_nenmatsu) %>% 
  rename(ref_date = date)
clust_param_date_tbl  

```

# Generate

-   This chunk executes parallel computations for each model execution date,
    using `clust_param_date_tbl` and `mkt_data_tbl` as inputs.

```{r}
if(!is_load){

  obj = create_global_objects_for_parallel_cluster_model(mkt_data_tbl)
  global_f = obj$f
  global_variables = obj$variables
  
  #グローバル環境でのオブジェクト定義
  environment(global_f) <- .GlobalEnv
  list2env(global_variables, envir = .GlobalEnv)
  environment(execute_parallel_computation_on_splits) <- .GlobalEnv
  
  clust_result_tbl = execute_parallel_computation_on_splits(
    clust_param_date_tbl, split_function, 
    add_libraries = c("mclust", "here", "R6", "rlang"), no_cores = no_cores) 
  rm(global_f, global_variables)
  if(is_prod){
    clust_handler$save_results(clust_trial, "clust_result_tbl", clust_result_tbl)  
  }
  
}else{
  
  clust_result_tbl = clust_handler$load_results(result_name =  "clust_result_tbl", trial = clust_trial)
}


```

```{r}
clust_result_tbl
```

```{r}
obj = clust_result_tbl %>% 
  filter(ref_date == max(ref_date)) %>% 
  pull(obj) %>% magrittr::extract2(1) 

```
